---
title: "Winning at work with R and the Tidyverse"
author: "Miles McBain"
date: "4 February 2017"
output: 
    ioslides_presentation:
        css: ./style.css
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(here)
COURSE_HOME <- here()
```

#Data Munging

## Dplyr
* `dplyr` is the data wrangling workhorse of the tidyverse.
* Prodvides functions, **verbs**, that can maipulate tibbles into the shape you need for analysis.
* Has many backends allowing dplyr code to work on data stored in SQL databases and big data clusters.
    - Works via translation to SQL. Keep an eye out for the SQL flavour in `dplyr`

##Basic Vocab
* `select()` columns from a tibble 
* `filter()` to rows matching a certain condition
* `mutate()` a tibble by changing or adding rows
* `arrange()` rows in order
* `group_by()` a variable
* `summarise()` data over a group using a function

##Select
```{r}
iris %>% 
    select(Sepal.Width, Species) %>%
    select(-Species) %>%
    head()
```
* What is the output of this code?
* Note the use of *bare* *unquoted* variable names.
    - No `iris$Species` or `iris[,"Species"]`
    - How does this even work?

##Filter

##Mutate


##Group by
`ungroup()`

#Create Copy

#Conditional Mutate?
`if_else()`

#SQL Backend