---
title: "Winning at work with R and the Tidyverse"
author: "Miles McBain"
date: "4 February 2017"
output: 
    ioslides_presentation:
        css: ./style.css
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(here)
COURSE_HOME <- here()
```

#Deep Dplyr

##Delving Further into dplyr
* `mutate` and `summarise` have conditional versions to make large scale wrangling a breeze.
* E.g. `mutate()` is part of a family that contains `mutate_at()`, `mutate_all()`, `mutate_if()`
    - Can do some pretty cool tricks!

##String/Factor Conversion    
```{r, eval=FALSE}
income_data <- read_csv(file.path(COURSE_HOME,"data/income.csv"))
    
income_data %>%
    mutate_if(.predicate = is.character, 
              .funs = as.factor)
```

##Any NA?
```{r, eval=TRUE}
airquality %>%
    summarise_all(.funs = anyNA)
```

##5-Number Summary for Groups
A 5 number summary for every month? Easy.
```{r}
airquality %>%
    group_by(Month) %>%
    summarise_at(.cols = vars(Ozone, Solar.R, Wind, Temp),
                 .funs = funs(q1 = quantile(., probs=0.25),
                              med = median,
                              mean = mean,
                              q3 = quantile(., probs = 0.75),
                              sd = sd 
                          ),
                 na.rm = TRUE
                 )
```

## Combining Data
* Sometimes you have multiple datasets you need to join together to create your tidy data set for analysis.
* Common scenarios where this arises:
    - Your data was extracted from a relational database (SQL): multiple tables
    - You need to append public data to your dataset. E.g weather, financial indicies.
    - Your data is published periodically. E.g. ABS stats, Financial reports etc.

## Joins
* `dplyr::bind_rows()`, `dplyr::bind_cols()` can work for trival cases where we have exactly the same number of rows or columns.
* For more complex cases we will need to consider a **join**.
* Before you can join two or more datasets you need to identify the **key** column(s).
    - These are columns that uniquely identify each row.
    
## Key example {.codefont}
What is the key?
```{r, eval=FALSE}
#install.packages(nycflights13)
library(nycflights13)
data("planes")
planes %>% View()
data("flights")
flights %>% View()
data("weather")
#weather %>% View()
```

## Join Example: Weather to Flights {.codefont}
Is departure delay associated with temperature?
```{r, eval=FALSE}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier, dep_delay)

flight_weather <-
  flights2 %>%
  left_join(weather, by = c("year", "month", "day", "hour", "origin"))

ggplot(flight_weather, aes(x=temp,y=dep_delay)) +
  geom_point(alpha=0.2)
```

##Types of joins
* We'll quickly walk though [the guide in R for Data Science](http://r4ds.had.co.nz/relational-data.html#understanding-joins).
    - This is a great resource for further reading.

##Issues arising when joining
* Missing values
* Duplicate values


