---
title: "Winning at work with R and the Tidyverse"
author: "Miles McBain"
date: "4 February 2017"
runtime: shiny_prerendered
output:
  bookdown::gitbook:
      dev: svglite
      css: css/style.css
      split_by: section
      config:
        toc:
          collapse: none
        download: [pdf, epub, mobi]
        edit: https://github.com/milesmcbain/tidycourse/edit/master/inst/examples/%s
        sharing:
          github: yes
          facebook: no
---

```{r setup, include=FALSE}
library(tutor)
library(tidyverse)
```

#Introduction
* What is the tidyverse?
* Tidy Data

##Prerequisites

#Tidystyle
[This is a good start](https://rpubs.com/yeedle/tidyguide)

##Pipes %>%

### Example
```{r, eval=FALSE}
mtcars %>% 
    mutate(pwr_2_weight = hp/wt) %>%
    arrange(pwr_2_weight) %>%
    head()
```

### Notes
* We call the `%>%` operator the 'pipe'.  
* Originally from the `magrittr` package but has been imported to the `tidyvsere`.
* `%>%` is an **infix operator**. This means it takes two arguments, one on the left and right.
* Pipe feeds the output of the last expression/function forward to the first input of the next funciton.
    - `.` can also control the input position if the next function

other infix operators:
```{r, eval=FALSE}
c(1,2) %*% t(c(3,4)) #vector/matrix multiplaction

"apple" %in% c("orange", "pear", "banana") #containment
```

### Example

```{r addition, exercise=TRUE}
1 + 1
```




<!--chapter:end:index.Rmd-->

